Joint Neighborhood Subgraphs Link Prediction

### 2017-06-29 paper.tex
Add files via upload

\section{Introduction}
A huge amount and variety of relational data are available in a wide range of fields. On the one hand, it provides an opportunity to build systems which automatically extract useful information. On the other hand, it raises challenges for machine learning and data scientists to design high performance predictive systems. A common data struture for representing relational data is graph whose nodes describe for entities and links depict their relations. We can easily see graph-based systems in different domains like biology, chemistry or networking. These systems require the data form used for their input as graphs. However, despite the availability of huge data, most graphs are only partially observed graphs. That means a number of actual links in each graph are absent or missing. Tackling this issue is a necessary condition to obtain high performance graph-based predictive systems. As a consequence, many link prediction methods have been proposed with the aim of recovering the missing links. A graph could be associtated with external information sources, so called "side information". Therefore, a link prediction method attempts to exploit the graph topology and side information to build a the model for predicting the presence or absence of link any between couple of nodes. An expected method is required to make use of both such kind of data sources. In this paper, we only focus on the topological information. Thus we the term link prediction methods to refer to topology-based link prediction methods.

Conerning link prediction methods, we can classify into two classes: unsupervised learning and supervised learning. Unsupervised learning methods compute the scores for node couples based on topological features and directly use these scores to consider non-observed links to be present or absent. Typical examples link prediction methods which fall into this class are Adamic-Adar \cite{adamic}, Katz \cite{katz}, Leicht-Home-Newman \cite{lhni}, etc. Supervised learning link prediction methods are based on a learning process which try to separate present links and absent links by their characteristics. A link prediction method via matrix factorization \cite{matrix-factorization} is for instance. This group of methods can also be further grouped into four subsets: feature-based models, graph regularization models, latent class models and latent feature models as presented in \cite{matrix-factorization}. For the performance, supervised methods tend to show better performance comparing to unsupervised ones in general. However, they normally face with the high compexity in term of computation and memory consumption.

@@

\section{Method}
In this section, we describe our proposed link prediction method which is based on the graph structure exploitation only.
\subsection{Definitions and notation}
We consider a graph $G=(V,E)$ where $V$ is the set of nodes and $E$ is the set of links. The set $E$ can be devided into a subset of observed links (O) and a subset of unobserved links (U). The problem is that the set O is not a complete set. Therefore, \textit{link prediction} is the task that aims at exploiting graph structure to rank links in $U$ from the most to the least probable to be considered as observed links in $O$.

@@

\subsection{Node labeling}
NSPDK takes labeled graphs as its input to proceed. Therefore, graph node labeling is a step needs to be done before employing NSPDK for graph similarity measure. Here, we propose to use node labeling fuctions $L$ that take into account node degrees for their labels such that $\ell(u) = \ell(v)$ if $|d(u) - d(v)|\leq \lambda$, in which $\lambda$ is a hyper parameter. It is due to the assumption that nodes with similar degrees tend to share common properties. The values of $\lambda$ depend on the histogram of graph node degree. We can use tuning in order to obtain suitable value for $\lambda$.
\subsection{Link representation}
In this section, we propose a method to represent links in the form of neighborhood subgraphs.

Let us consider two nodes $u$ and $v$ ($u, v \in V$), we first extract two neighborhood subgraphs with radius $R$ rooted at $u$ and $v$ to obtain $N_R(u)$ and $N_R(v)$, respectively. We then represent a link connecting $u$ and $v$ by a graph $G_{uv}$ formed by taking a union of these two achieved subgraphs: $G_{uv}= N_R(u) \cup N_R(v)$.
\subsection{Joint neighborhood subgraphs link prediction}
Given a graph $G(V,E)$ and a set of links $L = \lbrace(u,v)\ |\ (u,v)\in\ O\ or\ (u,v)\in U \rbrace$, our method consists five steps:
\begin{itemize}
	\item Node labeling: Nodes in $G$ are labeled using a defined node label function $\ell$.
	\item Link representing: Link in $L$ are represented as joint neighborhood subgraphs. As a result, we have a set of graphs $SG = \lbrace G_{uv}\ |\ (u,v) \in L \rbrace$
	\item Kernel computation: Obtained graph set $SG$ is used to compute a gram matrix, $K$, by using NSPDK. $K$ shows the similarities between graphs (links).
	\item Model configuration: $K$ is then fed into a kernel machine to build a model which tries to learn a function to separate observed links from non-observed ones in the sense of properties.
	\item Scoring and ranking: Now the model is used to return scores for coming links. These scores are used to rank links from the most to the least probable of being present in the graph.
\end{itemize}

@@

>>>>

\section{Introduction and related work}
A huge amount and variety of relational data are available in a wide range of fields. On the one hand, it provides an opportunity to build systems which automatically extract useful information. On the other hand, it raises challenges for machine learning and data scientists to design high performance predictive systems. A common data struture for representing relational data is graph whose nodes describe for entities and links depict their relations. We can easily see graph-based systems in different domains like biology, chemistry or networking. These systems require the data form used for their input as graphs. However, despite the availability of huge data, most graphs are only partially observed graphs. That means a number of actual links in each graph are absent or missing. Tackling this issue is a necessary condition to obtain high performance graph-based predictive systems. As a consequence, many link prediction methods have been proposed with the aim of recovering the missing links. A graph could be associtated with external information sources, so called "side information". Therefore, a link prediction method attempts to exploit the graph topology and side information to build a the model for predicting the presence or absence of link any between couple of nodes. An expected method is required to make use of both such kind of data sources. In this paper, we only focus on the topological information. Thus we the term link prediction methods to refer to topology-based link prediction methods. Conerning link prediction methods, we can classify into two classes: unsupervised learning and supervised learning.

Unsupervised learning methods compute the scores for node couples based on topological features and directly use these scores to consider non-observed links to be present or absent. In \cite{adamic}, a method, \textit{Adamic-Adar}, is proposed in which a score between two nodes is computed by a weighted sum over their common nodes. Nodes with smaller degree are weighted more comparing to the higher ones. Another method also considers node degree when computing proximity is presented. The probability of a existing link connecting two nodes is proportional to the product of their node degrees. Different from the method in \cite{adamic} which considers the common nodes, a link prediction method, \textit{Katz}, is introduced in \cite{katz} that takes into acount the number of common paths with different lengths between two nodes to measure the proximity between them. The shorter paths are given more weights, meanwhile the longer ones are weighted less. Another method which is a modification of \textit{Katz} notated as Leicht-Holme-Newman is presented in \cite{lhni}. In this method the similarity of a node couple is based on an assumption that two nodes are more similar if they have more similar intermediate nodes. Besides, in \cite{matrix-factorization} they also apply a raw singular value decomposition, SVD, on the adjacency matrix for link prediction. Generally, unsupervised learning methods takes the advantage from their complexity since they do not face with learning process. However, in many case, they do not show impressing results.

Supervised learning link prediction methods are based on a learning process which try to learn a curve that separates present links and absent links relied on their characteristics. This group of methods can also be further grouped into four subsets: feature-based models, graph regularization models, latent class models and latent feature models as presented in \cite{matrix-factorization}. In \cite{nonparametric}, a nonparametric latent feature models for link prediction is introduced in which latent features are extracted by based on an adaptation of Bayesain nonparametric approach. Besides, this method is able to concurrently learn number of latent features and assign which entities have these features. Another method related to latent features is presented in \cite{matrix-factorization}, they propose to use matrix factorization to extract latent features. This method can also take the output of unsupervised methods as its input to proceed. They show impressive results with different use of loss functions: square loss function, log loss function. Compared to unsupervised learning methods, supervised methods tend to show better performance in general. However, they normally face with the high compexity in term of computation and memory consumption.

@@

\section{Method}
In this section, we describe our proposed link prediction method which is based on the graph structure exploitation only.
\subsection{Definitions and notation}
We consider a graph $G=(V,E)$ where $V$ is the set of nodes and $E$ is the set of links. The set $E$ can be devided into a subset of observed links ($O$) and a subset of unobserved links ($U$). The problem is that the set $O$ is not a complete set. Therefore, \textit{link prediction} is the task that aims at exploiting graph structure to rank links in $U$ from the most to the least probable to be considered as observed links in $O$.

@@

\subsection{Node labeling}
NSPDK takes labeled graphs as its input to proceed. Therefore, graph node labeling is a step needs to be done before employing NSPDK for graph similarity measure. Here, we propose to use node labeling fuctions $L$ that take into account node degrees for their labels such that $\ell(u) = \ell(v)$ if $|d(u) - d(v)|\leq \lambda$, in which $\lambda$ is a hyper parameter. It is due to the assumption that nodes with similar degrees tend to share common properties. The values of $\lambda$ depend on the histogram of graph node degree. We can use tuning technique in order to obtain suitable value for $\lambda$.
\subsection{Link representation}
Most methods for link prediction attempt to computes similarities between nodes and consider to rank links. We desire to take links as examples to learn and predict. Therefore, we propse a method which model the context of two nodes to represent for link connecting them.

Let us consider two nodes $u$ and $v$ ($u, v \in V$), we first extract two neighborhood subgraphs with radius $R$ rooted at $u$ and $v$ to obtain $N_R(u)$ and $N_R(v)$, respectively. We then create an artificial node $w$. This node is connected to both $u$ and $v$. As a consequence, we achieve a single graph representing for a link connecting $u$ and $v$.
\subsection{Joint neighborhood subgraphs link prediction}
Given a graph $G(V,E)$ and a set of links $L = \lbrace(u,v)\ |\ (u,v)\in\ O\ or\ (u,v)\in U \rbrace$, our method consists five steps:
\begin{itemize}
	\item Node labeling: Nodes in $G$ are labeled using a defined node label function $\ell$.
	\item Link representing: Link in $L$ are represented as joint neighborhood subgraphs. As a result, we have a set of graphs $SG = \lbrace G_{uv}\ |\ (u,v) \in L \rbrace$
	\item Kernel computation: Obtained graph set $SG$ is used to compute a gram matrix, $K$, by using NSPDK. $K$ shows the similarities between graphs (links).
	\item Model configuration: $K$ is then fed into a kernel machine to build a model which tries to learn a function to separate observed links from non-observed ones in the sense of properties.
	\item Scoring and ranking: Now the model is used to return scores for coming links. These scores are used to rank links from the most to the least probable of being present in the graph.
\end{itemize}

@@

\bibitem{nonparametric} Miller, K., Michael, et al.: Nonparametric latent feature models for link prediction. Advances in neural information processing systems. 2009.
